### 2015-10-27 Documentation/fokkerplanck.tex
added documentation

\documentclass[]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[utf8]{inputenc} 
\usepackage{subcaption}
\usepackage{caption}
\usepackage{hyperref} 
%opening
\title{}
\author{}

\newcommand{\R}{\ensuremath{\mathbb{R}}} % commando zonder argumenten
\newcommand{\C}{\ensuremath{\mathbb{C}}}
\newcommand{\N}{\ensuremath{\mathbb{N}}}
\newcommand{\norm}[1]{\left\|#1\right\|} % commando's met argumenten
\newcommand{\pa}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\ppa}[2]{\frac{\partial^2 #1}{\partial #2^2}}
\newcommand{\dd}{\ensuremath{\mathrm{d}}}
\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

We compare solutions of partial differential equations of advection-diffusion-type

\begin{equation}
\label{fokkerplanck}
\pa{\rho(x,t)}{t} + \pa{(a(x) \rho(x,t))}{x} = D  \ppa{\rho(x,t)}{x} 
\end{equation}

generated in two ways:
\begin{itemize}
\item By explicitly solving eq. \eqref{fokkerplanck} using the discretization scheme\footnote{First-order upwind scheme for the advection part combined with the Forward-Time Central-Space-method for the difussion part}
\begin{equation}
\rho_i^{n+1} = \rho_i^n + \Delta t \left( \frac{D}{{\Delta x}^2} \left( \rho_{i+1}^{n} - 2 \rho_i^n + \rho_{i-1}^n \right)  - \frac{a(x)}{\Delta x} (\rho_i^n - \rho_{i-1}^n) \right)
\end{equation}
for the value of the probabilty density $\rho$ at position $i \Delta x$ and time $(n+1) \Delta t$.
\item By simulating an ensemble of $N$ particles evolving according to the corresponding stochastic differential equation  
\begin{equation}
    \dd X = a(X) \dd t + \sqrt{2D} \cdot \dd{W}.
\end{equation}
 The position $X^{n+1}$ at time $ (n+1) \Delta t$ of each  particle is simulated using the  Euler-Maruyama scheme
\begin{equation}
   X^{n + 1} = X^{n} + a(X^n) \Delta t +  \sqrt{2D \Delta t}\cdot \xi^n 
\end{equation}
with $\xi^n  \sim \mathcal{N} (0,1)$
\end{itemize}

\section{}

\end{document}