### 2016-01-15 Documentation/biblio.tex
newton_krylov

@article{avitabile14,
  author = "Avitabile, Daniele and Hoyle, Rebecca and Samaey, Giovanni",
  title = "Noise reduction in coarse bifurcation analysis of stochastic agent-based models: an example of consumer lock-in",
  journal = "SIAM Journal on Applied Dynamical Systems",
  volume = "13",
  number = "4",
  pages = "1583--1619",
  year = "2014",
  url = "https://lirias.kuleuven.be/handle/123456789/492023",
}

@article{watts98,
  author="Watts, Duncan J and Strogatz, Steven H",
  title= {\href{http://www.nature.com/nature/journal/v393/n6684/pdf/393440a0.pdf} {Collective dynamics of 'small-world' networks}},
  journal="Nature",
  volume="393",
  pages="440--442",
  year="1998"
}